### 2017-09-22 pres/proposal.tex
start pres

\documentclass[9pt]{beamer}

\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{tikz}

\usetheme{Boadilla}

\usefonttheme{professionalfonts}

\title[Short Title]{Title}
\subtitle{}
\author{Max Callaghan}
\institute[MCC]{\includegraphics[height=1cm,width=2cm]{/home/max/Pictures/MCC_Logo_RZ_rgb.jpg}}

\newtheorem*{remark}{}

\bibliographystyle{apalike}

\begin{document}

\begin{frame}
	\titlepage
\end{frame}

\addtobeamertemplate{frametitle}{}{%
	\begin{tikzpicture}[remember picture,overlay]
	\node[anchor=north east,yshift=2pt] at (current page.north east) {\includegraphics[height=0.8cm]{/home/max/Pictures/MCC_Logo_RZ_rgb.jpg}};
	\end{tikzpicture}}

\begin{frame}{Frame Title}

\end{frame}

\begin{frame}{Frame Title}
	\small
	\bibliography{/home/max/Documents/library/bibliography}
\end{frame}

\end{document}