### 2018-09-04 tables/ipcc_matching_sample.tex

\begin{tabular}{l l p{4.5cm} p{2.5cm} r}
\toprule
 AR &  WG &                                                                                                                                                                                                                        text &                                                   authors &  year \\
\midrule
 2 &  2 &  Landfill gas: working with Gaia. Biodeterioration Extracts no. 4, Energy Technology Support Unit, Harwell Laboratory, Oxfordshire, UK.  &  Richards, K.M. &  1989 \\
 1 &  1 &  Longwave cloud radiative forcing as determined from Nimbus-7 observations J Cltm , 2, 766 799  &  Ardanuy, P E , L L Stowe, A Gruber, M Weiss and C S Long &  1989 \\
 3 &  2 &  Climate change: overview and implications   for wildlife. In: Wildlife Responses to Climate Change: North American Case   Studies [Schneider, S.H. and T.L. Root (eds.)]. Island Press, Washington, DC,   USA, (in press). &  Root, T.L. and S.H. Schneider &  2001 \\
 2 &  3 &  The impact ol' global wanning on the United States: A survey of recent literature, mimco. Institute lor Interna-tional Economics, Washington, DC (April).  &  Cline, W.R. &  1993 \\
 3 &  2 &  Population-environment   relations at the forested frontier of Nepal. Applied Geography, 20, 221-242. &  Conway, D., K. Bhattarai, and N.R. Shrestha &  2000 \\
 3 &  2 &  The Cities Project. Australian Geological Survey Organisation,   Australia. Available online at http://www.agso.gov.au/geohazards/grm/cities2.html. &  AGSO &  1999 \\
 4 &  1 &  CLIMBER-2: A climate system model of intermediate complexity. Part I: Model description and performance for present climate.  &  Petoukhov, V., et al. &  2000 \\
 5 &  1 &  A skill-score based evaluation of simulated Australian climate. Australian Meteorol. Oceanogr &  Watterson, I., A. C. Hirst, and L. D. Rotstayn &  2013 \\
 5 &  1 &  Enhanced aerosol backscatter adjacent to tropical trade wind clouds revealed by satellite-based lidar. Geophys. Res &  Tackett, J. L., and L. Di Girolamo &  2009 \\
 5 &  3 &  Promoting long-term invest-ments by institutional investors. OECD Journal: Financial Market Trends 1, 145 – 164 &  Della Croce R, F Stewart, and J Yermo &  2011 \\
\bottomrule
\end{tabular}