### 2018-11-27 tables/growth_table.tex

\begin{tabular}{l p{1.8cm} p{1.8cm} p{1.8cm} p{1.8cm} p{1.8cm} p{1.8cm}} 
&\textbf{AR1} & \textbf{AR2} & \textbf{AR3} & \textbf{AR4} & \textbf{AR5} & \textbf{AR6}\\ \hline\textbf{Documents} &625 & 7623 & 16395 & 34510 & 117758 & 128266\\ 
\textbf{Words} &1380 & 12409 & 20453 & 32644 & 67064 & 74196\\ 
\textbf{New words} & change (296) & loss (552) & downscaling (197) & sres (217) & biochar (1752) & mmms (192)\\ & climate (262) & efficiency (515) & degreesc (145) & petm (95) & redd (1058) & c3n4 (132)\\ & model (168) & mol (439) & ncep (130) & amf (87) & cmip5 (656) & cop21 (107)\\ & effect (160) & ambient (417) & otcs (87) & sf5cf3 (81) & cmip3 (569) & cmip6 (104)\\ & co2 (156) & coal (404) & nee (87) & cwd (74) & wrf (334) & zika (75)\\ & atmospheric (152) & photosynthetic (393) & fco (80) & embankment (72) & mofs (288) & brgdgts (71)\\ & climatic (133) & concern (381) & hadcm2 (78) & aod (69) & sdm (283) & twitter (68)\\ & global (131) & chamber (353) & dtr (75) & clc (69) & gosat (281) & jing (66)
\end{tabular}