### 2018-11-27 tables/top_topics.tex

\begin{tabular}{p{1.9cm} p{3cm} p{7.5cm} r}
\toprule
                      title &                                                                            top words &                                                                                                                                                                                                                                                                top docs &   share \\
\midrule
      climat, chang, impact &          [climat, chang, impact, respons, futur, effect, shift, sensit, affect, may] &                                                                                                                                                  \parbox[t]{7.5cm}{Climate oscillations and changes over Russia; \\World Regionalization of Climate Change (1961-2010)} &  2.73\% \\
     soil, moistur, microbi &       [soil, moistur, microbi, organ, respir, content, miner, depth, matter, efflux] &                                                                                    \parbox[t]{7.5cm}{PARTITIONING OF SOIL RESPIRATION IN A FIRST ROTATION BEECH PLANTATION; \\Responses of soil respiration to N fertilization in a loamy soil under maize cultivation} &  2.73\% \\
       emiss, reduct, reduc &      [emiss, reduct, reduc, greenhous, factor, total, estim, inventori, nox, measur] &                                                                                                             \parbox[t]{7.5cm}{China's CH4 and CO2 emissions: Bottom-up estimation and comparative analysis; \\Monitoring total emissions from industrial installations} &  2.21\% \\
   carbon, dioxid, sequestr &   [carbon, dioxid, sequestr, sink, organ, cycl, storag, stock, terrestri, atmospher] &                             \parbox[t]{7.5cm}{Interpreting carbon-isotope excursions: carbonates and organic matter; \\PARTICULATE FLUXES OF CARBONATE AND ORGANIC-CARBON IN THE OCEAN - IS THE MARINE BIOLOGICAL-ACTIVITY WORKING AS A SINK OF THE ATMOSPHERIC CARBON} &  1.74\% \\
      temperatur, air, mean &  [temperatur, air, mean, surfac, minimum, maximum, daili, increas, effect, degreesc] &                                                                                                     \parbox[t]{7.5cm}{Observed changes in shallow soil temperatures in Northeast China, 1960-2007; \\Beyond the Mean: Biological Impacts of Cryptic Temperature Change} &  1.71\% \\
      record, dure, glacial &       [record, dure, glacial, reconstruct, last, period, holocen, event, late, core] &                                                           \parbox[t]{7.5cm}{HIGH-RESOLUTION CLIMATE RECORDS FROM THE NORTH-ATLANTIC DURING THE LAST INTERGLACIAL; \\HIGH-RESOLUTION CLIMATIC INFORMATION FROM SHORT FIRN CORES, WESTERN DRONNING MAUD LAND, ANTARCTICA} &   1.7\% \\
     speci, distribut, rang &          [speci, distribut, rang, rich, invas, nich, predict, extinct, shift, abund] &                               \parbox[t]{7.5cm}{Northward range extensions of some mesopelagic fishes in the Northeastern Atlantic; \\Natural occurrence and backwater infection of C-4 plants in the vegetation of the Yangtze hydropower Three Gorges Project region} &   1.7\% \\
 increas, concentr, decreas &   [increas, concentr, decreas, effect, atmospher, doc, result, organ, nutrient, may] &                                                          \parbox[t]{7.5cm}{TERRESTRIAL HIGHER-PLANT RESPONSE TO INCREASING ATMOSPHERIC [CO2] IN RELATION TO THE GLOBAL CARBON-CYCLE; \\Hydrological response to climate change in the Black Hills of South Dakota, USA} &  1.61\% \\
      forest, tropic, stand &       [forest, tropic, stand, deforest, disturb, stock, boreal, redd, harvest, wood] &  \parbox[t]{7.5cm}{Spatially explicit estimates and temporal changes of forest tree biomass in a typical department of forest management, Turkey; \\Analysis of the changes in forest ecosystem functions, structure and composition in the Black Sea region of Turkey} &  1.56\% \\
    energi, renew, consumpt &        [energi, renew, consumpt, effici, demand, save, sector, sourc, industri, use] &                                                                                                                                       \parbox[t]{7.5cm}{Energy issues and energy priorities; \\Energy efficiency and CO2 emissions in Swedish manufacturing industries} &  1.56\% \\
\bottomrule
\end{tabular}