### 2019-03-04 results.tex

\documentclass{article}
\usepackage[a4paper, total={7.5in, 10in}]{geometry}
\geometry{paperwidth=8in, paperheight=16383pt, left=1in, top=40pt, textwidth=6in, marginparsep=20pt, marginparwidth=100pt, textheight=16263pt, footskip=40pt}
\usepackage{graphicx}
\usepackage{url}
\usepackage{natbib}
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{lineno}
\usepackage{color}
%\usepackage{auto-pst-pdf}
\usepackage[colaction]{multicol}
\usepackage{caption}
\usepackage{svg}
\usepackage{authblk}
\usepackage{standalone}
\usepackage[section]{placeins}

\linespread{1.5}

\makeatletter
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
	\begin{flushleft}

		{\huge\textbf{\@title}}

		\bigskip

 		{\large\textbf{\@author}}

 		\bigskip

 		{\large{Draft current \@date}}

	\end{flushleft}\egroup
}
\makeatother

\newcommand{\multicollinenumbers}{
	\linenumbers
	\def\makeLineNumber{\docolaction
		{\makeLineNumberLeft}
		{}
		{\makeLineNumberRight}
		}
}

\newenvironment{figurehere}
	{\def\@captype{figure}}
	{}

% Title
\title{A Topic Model of Climate Change Literature}
\title{Words, words, words: Mapping the Matter of Climate Change Literature}
\title{A Topography of Climate Change Research - Results}
\author[1,2]{Max Callaghan}

\affil[1]{Mercator Research Institute on Global Commons and Climate Change, Torgauer Straße, 10829 Berlin, Germany}
\affil[2]{School of Earth and Environment, University of Leeds, Leeds LS2 9JT, United Kingdom}

\begin{document}
\maketitle

\section{Results}

\setcounter{totalnumber}{200}

\begin{table}[h]
	\scriptsize
	\input{tables/growth_table.tex}
	\caption{Growth in climate change literature}
	\label{growthtable}
\end{table}

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.5\linewidth]{plots/literature_size/volume_variety.pdf}
		%\captionof{figure}
		\caption{The volume and variety of literature on climate change has grown to unmanageable proportions. Each box represents a document-term matrix (unique documents x unique terms) of the abstracts written in each assessment period. The percentage of documents in which the average word occurs in is given in the key.
		}

		\label{growth}
	\end{center}
\end{figure}

\begin{figure}[h]
	\begin{center}
		%\includegraphics[width=1\linewidth]{tsne_results/plots/run_665_s_0_p200_double.pdf}
		\includegraphics[width=0.85\linewidth]{tsne_results/plots/run_1275_s_0_p100_double.png}
		\caption{A map of the literature on climate change. Document positions are obtained by reducing the topic scores to two dimensions via t-SNE Documents are coloured by working group citations (top) and web of science discipline category (bottom). See SI table for topic composition of each grid square}
		\label{map-double}
	\end{center}
\end{figure}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.85\linewidth]{plots/ipcc_representation/ipcc_rep_oecds_time.pdf}
		\caption{A map of the literature on climate change. Document positions are obtained by reducing the topic scores to two dimensions via t-SNE Documents are coloured by working group citations (top) and web of science discipline category (bottom). See SI table for topic composition of each grid square}
		\label{oecd_rep}
	\end{center}
\end{figure}

\begin{figure}[h!]
	\begin{center}
		%\includegraphics[width=1\linewidth]{tsne_results/plots/run_665_s_0_p200_double.pdf}
		\includegraphics[width=1\linewidth]{tsne_results/plots/run_1275_s_0_p100_evolution.png}
		\caption{A map of the literature on climate change. Document positions are obtained by reducing the topic scores to two dimensions via t-SNE Documents are coloured by working group citations (top) and web of science discipline category (bottom). See SI table for topic composition of each grid square}
		\label{map-double}
	\end{center}
\end{figure}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.85\linewidth]{plots/ipcc_representation/ipcc_rep_new1275_all.pdf}
		\caption{A map of the literature on climate change. Document positions are obtained by reducing the topic scores to two dimensions via t-SNE Documents are coloured by working group citations (top) and web of science discipline category (bottom). See SI table for topic composition of each grid square}
		\label{ipcc_rep}
	\end{center}
\end{figure}

\begin{table}[h!]
	\scriptsize
	\input{tables/top_topics.tex}
	\caption{Top 10 topics in climate change literature}
	\label{top-topics}
\end{table}

\end{document}