### 2019-05-08 tables/growth_table_large.tex
large table

\begin{tabular}{l p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm} p{2.8cm}} 
&\textbf{AR1} & \textbf{AR2} & \textbf{AR3} & \textbf{AR4} & \textbf{AR5} & \textbf{AR6}\\ \hline\textbf{Documents} &1167 & 8539 & 21716 & 38750 & 134413 & 201606\\ 
\textbf{Words} &2000 & 12480 & 23346 & 34637 & 71867 & 94746\\ 
\textbf{New words} & change (560) & oil (287) & downscaling (217) & sres (234) & biochar (1791) & mmms (313)\\ & climate (428) & deltac (283) & degreesc (187) & petm (95) & redd (1113) & cop21 (234)\\ & co2 (318) & whole (256) & ncep (130) & amf (88) & cmip5 (679) & c3n4 (214)\\ & climatic (289) & tax (254) & fco (107) & sf5cf3 (86) & cmip3 (587) & sdg (187)\\ & model (288) & landscape (249) & pfc (98) & clc (81) & mofs (299) & zika (182)\\ & atmospheric (281) & alternative (243) & otcs (98) & embankment (81) & sdm (297) & ndcs (168)\\ & effect (280) & availability (242) & dtr (95) & cwd (79) & mof (275) & indc (164)\\ & global (224) & life (239) & nee (89) & etm (75) & biochars (252) & indcs (134)
\end{tabular}