### 2019-05-22 poster/literatur.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Literaturverzeichnis

%\begin{übler Trick um die Orinalüberschrift zu unterdrücken}
\vspace*{1cm}
      \begin{center} \textbf{\Large Literatur} \end{center}
\vspace*{-4.5cm}
%\end{übler Trick}

\renewcommand{\baselinestretch}{1.05}\small\normalsize
\begin{thebibliography}{9}
\bibitem{drp} \textsc{C. Tam \& J. Webb,} ``Dispersion-Relation-Preserving Finite Difference Schemes for Computational Acoustics'', \textsl{J. Comput. Phys.} 107, 1993.

\bibitem{lddrk} \textsc{F. Q. Hu, M. Y. Hussani \& J. L. Manthey,}  ``Low-Dissipation and -Dispersion Runge-Kutta Schemes for Computational Acoustics'',  \textsl{J. Comput. Phys.} 124(1), 1996.

\bibitem{shortwave} \textsc{C.K.W. Tam, J.C. Webb \& Z. Dong,} ``A study of the short wave components in computational acoustics'', \textsl{J. Comput. Acoust.} 1, 1993.

\bibitem{hamilton} \textsc{Hamilton \& Blackstock}, \textit{Nonlinear Acoustics}, Academic Press, 1997.

\bibitem{blackstock} \textsc{Blackstock \& Hamilton}, \textit{Nonlinear Acoustics}, Academic Press, 1997.
\end{thebibliography}