### 2019-07-10 main.tex

\documentclass{article}
\usepackage[a4paper, total={180mm, 260mm}]{geometry}
\usepackage{graphicx}
\usepackage{url}
\usepackage{natbib}
\usepackage{todonotes}
\usepackage{booktabs}
\usepackage{lineno}
\usepackage{color}
%\usepackage{auto-pst-pdf}
\usepackage[colaction]{multicol}
\usepackage{caption}
\usepackage{svg}
\usepackage{authblk}
\usepackage{standalone}

\usepackage[section]{placeins}

\linespread{1.5}

\makeatletter
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
	\begin{flushleft}

		{\huge\textbf{\@title}}

		\bigskip

		{\large\textbf{\@author}}

		\bigskip

		{\large{Draft current \@date}}

	\end{flushleft}\egroup
}
\makeatother

\newcommand{\multicollinenumbers}{
	\linenumbers
	\def\makeLineNumber{\docolaction
		{\makeLineNumberLeft}
		{}
		{\makeLineNumberRight}
	}
}

\newenvironment{figurehere}
{\def\@captype{figure}}
{}

% Title
\title{A Topic Model of Climate Change Literature}
\title{Words, words, words: Mapping the Matter of Climate Change Literature}
\title{A Topography of Climate Change Research}
\author[1,2]{Max Callaghan}
\author[1,2]{Jan Minx}
\author[2]{Piers M. Forster}

\affil[1]{Mercator Research Institute on Global Commons and Climate Change, Torgauer Straße, 10829 Berlin, Germany}
\affil[2]{Priestley International Centre for Climate, University of Leeds, Leeds LS2 9JT, United Kingdom}

\begin{document}
	\maketitle

	\begin{linenumbers}

		\noindent\textbf{\input{abstract.tex}}

		\bigskip

		\noindent We live in an age of ``Big Literature'' 
		\cite{Nunez-Mir2016, Minx2017l}, where the science of climate change is expanding exponentially \cite{Grieneisen2011, Haunschild2016}. In the five years since the publication of the last IPCC assessment report \cite{IPCC2014c}, 201,606 papers were published in the Web of Science (WoS) (see Table \ref{tab}). This is almost as much as the 204,585 papers published during the first five assessment periods, a period of nearly 30 years. A total of around 350,000 new publications can be expected for the current sixth assessment cycle of the IPCC, based on current growth patterns. Moreover, the literature has also become more diverse. This is reflected in the expansion of the literature's vocabulary - from 2,000 unique words in the first assessment period to 94,746 words so far in the sixth - indicating that the field has incorporated new content. For example, the zika virus, which was mentioned in 182 articles from 2014-2018, had never before been discussed in the titles or abstracts of articles relating to climate change. Yet it has emerged as a topic of high relevance: the incidence of the virus, the outbreak of which in Brazil in 2016 was declared a public health emergency by the WHO, is set to increase under rising global temperatures \cite{Rao2019}. Similar rapid emergence patterns can be seen for [Intended] Nationally Determined Contributions ([I]NDCs) and Sustainable Development Goals (SDGs) in AR6, and Biochar and REDD (The United Nations Collaborative Programme on Reducing Emissions from Deforestation and Forest Degradation in Developing Countries) in AR5, among others.

		\begin{table}[htp]
			{\scriptsize
				\input{tables/growth_table.tex}}
			\caption{Growth of Literature on Climate Change. A glossary of acronyms is provided in SI}
			\label{tab}
		\end{table}

		Big literature poses at least three challenges for scientific policy advice and science itself: \emph{First}, established procedures in scientific assessments like those conducted by the IPCC fail to address the exploding literature base. For example, the ratio of studies cited in IPCC reports to the number or relevant studies has declined from 60\% to 20\%  \cite{Minx2017l}, posing a rapidly growing risk of selection bias. More generally, the provision of comprehensive, objective, open and transparent assessments of the available scientific literature, as defined in the principles governing IPCC work \cite{IPCC2013}, is no longer possible for authors or author teams by traditional means. Machine reading and learning methods as well as other data science applications are required to enable an understanding of the field of climate change research at scale (REFs). \emph{Second}, evidence synthesis - the enterprise of reviewing the literature based on a formal and systematic set of methods \cite{Chalmers2002} - becomes increasingly important for aggregating and consolidating the rapidly emerging knowledge and enabling scientific assessments to do their job. Yet, traditional methods of evidence synthesis themselves are pushed to their limits by the large amount of scientific publications. Whilst the field of evidence synthesis technology that tries to streamline human tasks through machine learning at the different stages of the review process is still in its infancy \cite{Beller2018}. \emph{Finally}, overwhelming amounts of literature may be a major reasons why assessments of scientific assessments \cite{Bjurström2011, Hulme2010, Victor2015, Corbera2016} do not offer robust quantifications, for example, for claims about the relationship between report citations and the underlying literature. This may substantially affect findings.

		This study uses topic modelling \cite{Blei2010} to map out the vast body of evidence on climate change. Topic modelling is an unsupervised machine-learning technique, where patterns of word co-occurrences in documents are used to learn a set of topics which can be used to describe the corpus. The word topic derives from the Greek word for place (topos), and by \textit{situating} the documents in a reduced-form projection of their thematic content (see Figure \ref{oecd_topic_map}), we create a \textit{topographic map} of the literature on climate change. Such a systematic engagement with the thematic content of the climate science is missing from the literature so far.%To our knowledge, this is the first
		We apply this map in a second step to understand how the IPCC reports have represented the available climate change literature and re-evaluate claims of bias based on a more comprehensive understanding of the available climate science, and enrich the discussion of representation in the literature to include topics as well as disciplines.

		\subsection*{Mapping out the landscape of climate change literature}

		Figure \ref{oecd_topic_map} shows a \emph{thematic} or \emph{topopographic map} of the 400,000 publications on climate change in our dataset with a total number of 140 topics. The number of topics must be defined exogenously, but the results are robust to different specifications (see methods for examples and a thorough explanation of model selection). Using non-negative matrix factoriziation \cite{Lee1999}, the topics are machine-learned from the papers' abstracts (see methods for details).

		\begin{figure}[htp]
			\begin{center}
				\includegraphics[width=180mm]{plots_pub/all_topic_words_oecds.png}
				\caption{A map of the literature on climate change. Document positions are obtained by reducing the topic scores to two dimensions via t-SNE Documents are coloured by web of science discipline category. Topic labels are placed in the center of each of the large clusters of documents associated with each topic. }
				\label{oecd_topic_map}
			\end{center}
		\end{figure}

		The map shown covers a broad range of topics that occur in meaningful clusters. In general, topics related to climate science and impacts are in the West, while solution-oriented topics are in the East. More fine-grained research areas can also be distinguished. For example, publications related to urban infrastructure (buildings, energy, cement, waste) are located in the East (), physical climate impacts such as sea-level rise (), heat waves (), droughts () or crop yield (losses) () are in the South-West and energy systems are in North-East (). There are larger groups of documents at the fringes of the map that relate mainly to one or two specific topics such as biochar, corals, soil organic carbon, clouds and aerosols, diseases, membranes, absorption or adsorption or CO2 storage. Interestingly, scenarios feature centrally in the map, at the interface between different scientific communities. This corresponds to their integrative nature in IPCC reports (REF). This depiction of the thematic structure of the literature, and its ordering and categorising thereby could be useful for individual communities or for climate change assessments \footnote{We include the list of topic-document combinations in the supplementary materials}.

		The disciplinary composition of this research topography indicated by the different colours in Figure \ref{oecd_topic_map} highlights the dominance of natural sciences in climate change research. More than 60\% of the literature is published in natural science journals. Similarly, only in X out of 140 topics do publications from another field make up the majority of documents, in Y cases the contribution of publications in natural science journals make more than 50\%. We calculate disciplinary entropy of topics as a measure of their degree of interdisciplinarity (see Figure \ref{dis-entropy} and methods for details). While research on health, food, or policy comes from a range of disciplines, research on ice and oceans comes almost exclusively from the natural sciences).

		\begin{figure}
			\begin{center}
				\includegraphics[width=180mm]{plots_pub/topic_evolution_4.png}
				\caption{Evolution of the landscape of climate change literature. In each period, the 10 fastest growing topics are labelled. Where documents could be matched to IPCC citations, they are coloured by the working group citing them.}
				\label{evolution-map}
			\end{center}
		\end{figure}

		Finally, the topography shows the thematic evolution of the literature, with topics exhibiting distinct patterns of growth. Fast-growing topics in the last three assessment periods have included, among others, coral, risks, adaptation, hydrogen, buildings, CO2 removal, networks and biochar. Biochar is particularly remarkable in that the sizeable literature which emerged in AR5 was completely absent from the climate change literature beforehand. 
		The identification of new topics as they emerge, particularly as these are discovered without prior knowledge of the literature, can help researchers and assesment-makers to keep abreast of a quickly evolving field.

		\subsection*{Research representation in IPCC reports}
		We apply our topic map to understand the representation of science in IPCC assesssments and how it manages to respond to demands for more solution-oriented knowledge \cite{Kowarsch2017}. Several studies have identified, made, or repeated claims of a disciplinary bias of IPCC assessments towards the natural sciences, and within the social sciences towards economics \cite{Bjurström2011, Victor2015, Hulme2010, Corbera2016}. Where these claims were based on an analysis of IPCC citations \cite{Bjurström2011}, they fail to assess this claim against a reasonable, measurable benchmark. We argue here that the composition of the climate change literature as a whole provides such a benchmark, in view of the organisation's mandate to provide ``comprehensive, objective, open and transparent'' assessment of the available science \cite{IPCC2013}. Our comprehensive database allows us to study representation with such a benchmark, and over time rather than for single assessment cycles.

		Figure \ref{oecd_rep}.a shows that the social sciences were indeed under-represented in the third assessment report, but by the fifth assessment report were over-represented. Likewise, other social sciences than economics have become better represented since AR3  (see figure \ref{subfield}f) with social \& economic geography (4.3\% of the literature), political science (1.0\%), and sociology (0.8\%) showing improved representation in AR5 compared to AR3, and social and economic geography, political science, and other social sciences better represented than economics when taking into account the distribution of studies in the literature.

		This challenges what we think we know about the IPCC. The social sciences, by now, are actually the best represented field, with a share in the literature cited by IPCC reports [1.5] times as high as in the literature at large.  On the other hand the Agricultural Sciences and Engineering \& Technology have been consistently under-represented, with [2.5] and [3.5] times the share of studies in the wider literature than in the literature cited by the IPCC in AR5 respectively.

		The topography allows us to delve deeper into the subject matter that receives more or less attention in the IPCC. Figures \ref{oecd_rep}b and \ref{oecd_rep}c plot the representation of the topics shown in the map. Figure \ref{oecd_rep}c shows that topics more commonly cited by IPCC working group I are older and largely better represented in IPCC reports. These topics, for example \textbf{ozone}, \textbf{oceans}, \textbf{clouds}, \textbf{aerosols} and \textbf{sea levels}, are older, predominantly cited by WG I, and well cited by IPCC reports. They make up some of the core topics of the physical science of climate change.

		The topics in the lower right of the graph are the most pertinent to the question of whether the IPCC is well representing knowledge on climate change. These topics are newer and until now have been under-represented in IPCC reports. They are the topics that could be seen as potential gaps in the IPCC's coverage of the science. Because they are new areas of knowledge, they may be highly salient in a periodic assessment process. These topics are primarily in working group III, on mitigation (see methods section [] for a discussion of the categorisation of topics, including CLC, adsorption and hydrogen, which may more properly be of relevance to WGIII).

		The difference between these under-represented new topics and other new topics that are better represented is intriguing. This difference is visible in figure \ref{evolution-map}, where In AR5, the clusters of documents around the \textbf{adsorption}, \textbf{buildings}, and \textbf{biochar} topics contain few IPCC citations, whereas the clusters around \textbf{food}, \textbf{health}, \textbf{adaptation}, and \textbf{GHGs} contain more. As shown in figure \ref{oecd_rep}c, adsorption buildings and biochar are [4, 3.5 and 3.8] times more prevalent in the literature than in IPCC citations, while food is [0.3] times more prevalent in the literature and health and adaptation are [0.1] and [2,5] times more prevalent in IPCC citations respectively. The IPCC, then, has been better at integrating new knowledge from these topics, and in general better at integrating new knowledge from WG II  than WG III topics.

		Further, within WG III topics, those that are well represented contain a greater proportion of social science research (figure \ref{oecd_rep}b). The topics \textbf{countries}, \textbf{policy}, and \textbf{prices} are close to a proportional representation and are made up of around 30\% social science research. \textbf{Waste}, \textbf{biochar}, \textbf{cement} and \textbf{coal}, are more than 3 times more prevalent in the wider literature than in the literature cited by the IPCC, and are made up of around 5\% social science research. This pattern is not visible in other working groups (see Figure \ref{socsci-wgs}), and complicates the perception of the under-representation of the social sciences.

		Recalling policymakers' demands for more solution-oriented assessments \cite{Kowarsch2017}, we could also interpret the topics that are newer and under-represented as ``solutions-relevant''. Many deal with negative emissions, or with mitigation (or to a lesser extent adaptation) options in the transport, buildings and power sectors. All are arguably related to climate solutions. However, while policymakers' demands for solutions-oriented knowledge were rather about policy options, these under-represented new topics deal with more technical solutions and are found rather in technical disciplines within engineering \& technology and the agricultural sciences.

		\begin{figure}[htp]
			\begin{center}
				\includegraphics[width=180mm]{plots_pub/big_panel_representation.pdf}
				\caption{Representation in IPCC reports: \textbf{a)} by discipline, \textbf{b)} by social science proportion of WG 3 topics, \textbf{c)} and novelty of all topics, where topics in the highest and lowest 10\% of either axis are labelled. Topics are coloured according to the working group from which they receive the most citations. Representation is the share of the subset of documents being cited by the IPCC divided by the share of the subset in the whole literature. The log is taken so that 0 is equal to perfectly proportional representation, and -1 and 1 are equally under and over represented.}
				\label{oecd_rep}
			\end{center}
		\end{figure}

		\subsection*{Machine-learning for climate change assessments}%, evidence synthesis and science of science} % needs to be reviewed
		Keeping track of developments in the scientific literature is becoming increasingly difficult in the age of big literature. This is becoming a central challenge for global environmental assessments such as those conducted by the IPCC \cite{Minx2017l, Kowarsch2017}, as well as for science of science research (\cite{Zeng2017}), the provision of thorough and comprehensive reviews in specific fields, and the efficient allocation of research funding where it is needed. Ultimately it is a challenge for sound, evidence-based scientific policy advice (\cite{White2019}). Systematic maps and evidence gap maps (\cite{James2016}, \cite{McKinnon2015}) provide crucial tools for tracing the literature, but current approaches are constrained by available resources to answering very specific questions, or are limited by a lack of suitable literature.  On the other hand, current scientometric and research weaving approaches (\cite{Nakagawa2019}) can capture a broader landscape of research, but are too unspecific with regard to the thematic content of literature.

		Elsewhere, we show how machine learning can accelerate the provision of traditional maps of evidence with thousands rather than hundreds of data points (\cite{Lamb2018d}). Here we extend this analysis to vast bodies of literature with hundreds of thousands of publications and show how ``topopgraphic maps'' provide an efficient visualisation of these.

		We acknowledge that the climate change topography presented here is not fully comprehensive. The WoS  covers a substantive, yet by no means exhaustive set of publications. Grey literature, some smaller journals and books, not to mention indigenous knowledge \cite{Ford2016b}, are not covered. Similarly, our query (see methods) requires that the keywords, title or abstract of a publication mentions climate change or a synonym or related term. Whilst this is a reasonable system boundary definition, and follows \cite{Grieneisen2011} exactly, we acknowledge the existence of directly relevant studies for climate change research - for example on energy technologies, energy policies or - that may not meet our criteria and remain outside the scope of our analysis. Yet our decisions are driven by data availability and requirements for a clean, climate-changed focused dataset. Given the large size of our sample (400,000+ publications), we argue that our topography provides a reasonable representation of the research landscape on climate change. Furthermore, unless IPCC citation patterns systematically differ in the not included parts of the literature, the main conclusions of this study are unaffected.

		In addition, our analysis is based on a topic model of publication abstracts. Though it is reasonable to assume that abstracts reflect the themes of the papers they describe, this depends on the quality of the abstracts, and thematic features of the underlying texts may not be adequately represented.  In the absence of comprehensive full-text access and remaining, non-trivial challenges in text extraction from PDFs, we argue that abstracts provide a reasonable unit of analysis.

		We apply our map of the climate change literature to understand the representativeness of IPCC citations and argue that answering related questions requires a clear benchmark. We show that applying this benchmark challenges and confirms some prominent findings around disciplinary representation in and solution-orientation of IPCC assessments.

		While proportional representation of all topics and all disciplines may not be desirable, the analysis presented here can help the IPCC to make quantitatively informed decisions about what to represent. The claim that the IPCC needs to incorporate more social science research may indeed be well-founded, but, in light of the fact shown here that the IPCC over-represents the social science literature that is published, this would imply increases in funding for climate change research in the social sciences. While some might argue that engineering or agricultural research on solutions related topics may be too technical and less relevant to IPCC assessments, in light of the size of the literature in these areas relative to that which is cited by the IPCC it would follow that more social science research needs to be carried out that deals with these topics and produces assessment-ready knowledge.

		We highlight that there are a variety of other immediate applications for such comprehensive topic maps and their rich, underlying data. Rapid, comprehensive maps could make a contribution to making climate change assessment processes in various places more efficient. For example, they could inform the scoping of scientific assessment reports to make sure that reports keep abreast of developments in the literature and remain policy relevant. What's more, they could help in the efficient identification of the best subject area experts across regions, sectors and genders (REFs) and directly increase the transparency with regard to the availability and selection of citations. Beyond the organisation of global environmental assessments, such maps can inform research funding decisions, and help individual researchers identi Future research can explore how further methods, and more comprehensive data, can help drive forward these applications.

		While the size and variety of the literature on climate change presents challenges to global environmental assessments, it represents a wealth of knowledge on vital climate solutions. By shedding light on new and under-reported subjects and by providing quantifications of what we think we know about how assessments and their relationship to the literature, computational mapping can help ensure that this knowledge is made use of efficiently.  Topographic maps  can become part of a toolbox which helps the IPCC continue to be able to fulfil its mandate of providing comprehensive, objective and transparent assessments of an expanding literature.

	\end{linenumbers}

	\appendix

	%\listoffigures
	\linespread{1}
	%\bibliography{Mendeley}
	\begin{thebibliography}{10}

		\bibitem{Bjurström2011}
		Andreas Bjurstr{\"{o}}m and Merritt Polk.
		\newblock {Physical and economic bias in climate change research: A
			scientometric study of IPCC Third Assessment Report}.
		\newblock {\em Climatic Change}, 108(1):1--22, 2011.

		\bibitem{Nunez-Mir2016}
		Gabriela~C Nunez-Mir, Basil~V Iannone, Bryan~C Pijanowski, Ningning Kong,
		Songlin Fei, and Richard Fitzjohn.
		\newblock {Automated content analysis: addressing the big literature challenge
			in ecology and evolution}.
		\newblock {\em Methods in Ecology and Evolution}, 7(11):1262--1272, 2016.

		\bibitem{Grieneisen2011}
		Michael Grieneisen and Minghua Zhang.
		\newblock {The Current Status of Climate Change Research}.
		\newblock {\em Nature Climate Change}, 1:72--73, 2011.

		\bibitem{Haunschild2016}
		Robin Haunschild, Lutz Bornmann, and Werner Marx.
		\newblock {Climate Change Research in View of Bibliometrics}.
		\newblock {\em PLoS ONE}, 11(7):1--19, 2016.

		\bibitem{IPCC2014c}
		IPCC.
		\newblock {\em {Climate Change 2014: Synthesis Report. Contribution of Working
				Groups I, II and III to the Fifth Assessment Report of the Intergovernmental
				Panel on Climate Change}}.
		\newblock IPCC, Geneva, Switzerland, 2014.

		\bibitem{Rao2019}
		V.~Brahmananda Rao, K.~Maneesha, Panangipalli Sravya, Sergio~H. Franchito,
		Hariprasad Dasari, and Manoel~A. Gan.
		\newblock {Future increase in extreme El Nino events under greenhouse warming
			increases Zika virus incidence in South America}.
		\newblock {\em npj Climate and Atmospheric Science}, 2(1):2--8, 2019.

		\bibitem{Minx2017l}
		Jan~C. Minx, Max Callaghan, William~F. Lamb, Jennifer Garard, and Ottmar
		Edenhofer.
		\newblock {Learning about climate change solutions in the IPCC and beyond}.
		\newblock {\em Environmental Science {\&} Policy}, 2017.

		\bibitem{IPCC2013}
		IPCC.
		\newblock {Principles governing IPCC work}, 2013.

		\bibitem{Chalmers2002}
		Iain Chalmers, Larry~V Hedges, and Harris Cooper.
		\newblock {A Brief History of Research Synthesis}.
		\newblock {\em Evaluation {\&} The Health Professions}, 25(1):12--37, 2002.

		\bibitem{Beller2018}
		Elaine Beller, Justin Clark, Guy Tsafnat, Clive Adams, Heinz Diehl, Hans Lund,
		Mourad Ouzzani, Kristina Thayer, James Thomas, Tari Turner, Jun Xia, Karen
		Robinson, Paul Glasziou, and {On behalf of the founding members of the ICASR
			group}.
		\newblock {Making progress with the automation of systematic reviews:
			principles of the International Collaboration for the Automation of
			Systematic Reviews (ICASR)}.
		\newblock {\em Systematic Reviews}, 7(1):1--7, 2018.

		\bibitem{Hulme2010}
		Mike Hulme and Martin Mahony.
		\newblock {Climate change: What do we know about the IPCC?}
		\newblock {\em Progress in Physical Geography}, 34(5):705--718, 2010.

		\bibitem{Victor2015}
		{David G. Victor}.
		\newblock {Embed the social sciences in climate policy - David Victor}.
		\newblock {\em Nature}, 520:7--9, 2015.

		\bibitem{Corbera2016}
		Esteve Corbera, Laura Calvet-Mir, Hannah Hughes, and Matthew Paterson.
		\newblock {Patterns of authorship in the IPCC Working Group III report}.
		\newblock {\em Nature Climate Change}, 6(1):94--99, 2016.

		\bibitem{Blei2010}
		David Blei, Lawrence Carin, and David Dunson.
		\newblock {Probabilistic topic models}.
		\newblock {\em IEEE Signal Processing Magazine}, 2010.

		\bibitem{Lee1999}
		D~D Lee and H~S Seung.
		\newblock {Learning the parts of objects by non-negative matrix factorization.}
		\newblock {\em Nature}, 401(6755):788--791, 1999.

		\bibitem{Kowarsch2017}
		Martin Kowarsch, Jason Jabbour, Christian Flachsland, Marcel T.~J. Kok, Robert
		Watson, Peter~M. Haas, Jan~C. Minx, Joseph Alcamo, Jennifer Garard, Pauline
		Riousset, L{\'{a}}szl{\'{o}} Pint{\'{e}}r, Cameron Langford, Yulia Yamineva,
		Christoph von Stechow, Jessica O'Reilly, and Ottmar Edenhofer.
		\newblock {A road map for global environmental assessments}.
		\newblock {\em Nature Climate Change}, 7(6):379--382, 2017.

		\bibitem{Zeng2017}
		An~Zeng, Zhesi Shen, Jianlin Zhou, Jinshan Wu, Ying Fan, Yougui Wang, and
		H.~Eugene Stanley.
		\newblock {The science of science: From the perspective of complex systems}.
		\newblock {\em Physics Reports}, 714-715:1--73, 2017.

		\bibitem{White2019}
		Howard White.
		\newblock {The twenty-first century experimenting society: the four waves of
			the evidence revolution}.
		\newblock {\em Palgrave Communications}, 5(1):1--7, 2019.

		\bibitem{James2016}
		Katy~L. James, Nicola~P. Randall, and Neal~R. Haddaway.
		\newblock {A methodology for systematic mapping in environmental sciences}.
		\newblock {\em Environmental Evidence}, 5(1):1--13, 2016.

		\bibitem{McKinnon2015}
		Madeleine~C. McKinnon, Samantha~H. Cheng, Ruth Garside, Yuta~J. Masuda, and
		Daniel~C. Miller.
		\newblock {Sustainability: Map the evidence}.
		\newblock {\em Nature}, 528(7581):185--187, 2015.

		\bibitem{Nakagawa2019}
		Shinichi Nakagawa, Gihan Samarasinghe, Neal~R. Haddaway, Martin~J. Westgate,
		Rose~E. O'Dea, Daniel~W.A. Noble, and Malgorzata Lagisz.
		\newblock {Research Weaving: Visualizing the Future of Research Synthesis}.
		\newblock {\em Trends in Ecology and Evolution}, 34(3):224--238, 2019.

		\bibitem{Lamb2018d}
		William~F. Lamb, Felix Creutzig, Max~W. Callaghan, and Jan~C. Minx.
		\newblock {Learning about urban climate solutions}.
		\newblock {\em Nature Climate Change}, 9(4):279--287, 2018.

		\bibitem{Ford2016b}
		James~D. Ford, Laura Cameron, Jennifer Rubis, Michelle Maillet, Douglas
		Nakashima, Ashlee~Cunsolo Willox, and Tristan Pearce.
		\newblock {Including indigenous knowledge and experience in IPCC assessment
			reports}.
		\newblock {\em Nature Climate Change}, 6(4):349--353, 2016.

	\end{thebibliography}
	\bibliographystyle{unsrt}

	\include{methods}

\end{document}