### 2019-08-21 manuscript/abstract.tex

\documentclass{article}

\begin{document}
   The massive expansion of scientific literature on climate change poses challenges for global environmental assessments and how we
understand them. 
Big data and machine learning can help us deal 
with the large collections of text represented by scientific fields.
Such methods help make the production of assessments
more tractable, and give us better insights about how past assessments  have engaged with the literature as it has evolved.
We use topic modelling to identify the thematic structure and draw a comprehensive topic map, or topography, of over 400,000 scientific publications from the Web of Science on climate change. 
We update current knowledge on the IPCC, showing that the social sciences are in fact over-represented in recent assessment reports, and that
technical, solutions-relevant knowledge - especially in the agricultural and engineering sciences - are under-represented.
We point to a variety of other applications of such maps, and our findings have direct implications for addressing growing demands for more solution-oriented climate change assessments that are also more firmly rooted in the social sciences.
\end{document}