### 2019-10-01 tables/ipcc_matching_sample.tex
new plots after revisions

\begin{tabular}{l l p{4.5cm} p{2.5cm} r}
\toprule
 AR &  WG &                                                                                                                                                                           text &                                                            authors &  year \\
\midrule
 3 &  1 &  The IIASA Database for Mean Monthly   Values of Temperature, Precipitation and Cloudiness on a Global Terrestrial   Grid. IIASA Report RR-91-18, Laxenburg, 63 pp. &  Leemans, R. and W. Cramer &  1991 \\
 2 &  2 &  Rischi connessi con la dinámica glaciale nelle Alpi ItaUane. Geografía Física e Dinámica Quatemaría, 15, 85-99.  &  Dutto, F. and G. Mortara &  1992 \\
 3 &  1 &  Specification   of eddy transfer coefficients in coarse-resolution ocean circulation models.   J. Phys. Oceanogr., 27, 381-402. &  Visbeck, M., J. Marshall, T. Haine and M. Spall &  1997 \\
 3 &  2 &  A koala (Phascolartos   cinereus Goldfuss) population crash during drought and heatwave conditions in   south-western Queensland. Australian Journal of Ecology, 13, 451-461. &  Gordon, G., A.S. Brown, and T. Pulsford &  1988 \\
 4 &  2 &  Climate change and land resources of Russia. &  Isaev, A., V. Stolbovoi, V. Kotlyakov, S. Nilsson and I. McCallum &  2004 \\
 4 &  3 &  Generation technology choices: near and long term. US DOE/EPRI Annual Energy Outlook Conference, Washington D.C., www.eia.doe.gov/oiaf/archive/aeo04/conf/pdf/gehl.pdf  &  Gehl, S &  2004 \\
 4 &  3 &  Climate Change and the European Countryside: Impacts on Land Management and Response Strategies &  Viner, D., M. Sayer, M. Uyarra, and N. Hodgson &  2006 \\
 5 &  2 &  Adaptation to climate change in the context ofsustainable development and equity &  Smit, B. and O. Pilifosova &  2001 \\
 5 &  3 &  Copenhagen City of Cyclists: Bicycle Account 2010. City of Copenha-gen, The Technical and Environmental Administration. Available at: http: / / www. cycling-embassy &  COP &  2010 \\
 5 &  3 &  Uncertainty analysis of the IEA / ORAU CO2 emissions model. The Energy Journal 8, 1 – 29 &  Reilly J M, J A Edmonds, R H Gardner, and A L Brenkert &  1987 \\
\bottomrule
\end{tabular}