### 2019-12-13 manuscript/coverletter.tex

\documentclass{letter}
\usepackage{letterbib}
\usepackage{natbib}
\usepackage{geometry}

\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{url}
\usepackage{xifthen}

\usepackage{tikzpagenodes}
\geometry{
	a4paper,
	left=20mm,
	right=60mm,
	marginparwidth=50mm,
	top=30mm,
}

\def\myyshifttop{0}
\def\mypagetop{0}

\newcommand{\marpartop}[1]% content, color
{   \begin{tikzpicture}[remember picture, overlay]
	\ifthenelse{\thepage=\mypagetop}{}{\xdef\myyshifttop{0}}        
	\xdef\mypagetop{\thepage}
	\node[below right, yshift=\myyshifttop, text width=\marginparwidth-4pt, inner sep=2pt] (tempnode) at (current page marginpar area.north west) {#1};
	\path (current page marginpar area.north west);
	\pgfgetlastxy{\tempxone}{\tempyone}
	\path (tempnode.south west);
	\pgfgetlastxy{\tempxtwo}{\tempytwo}
	\pgfmathsetmacro{\diffy}{(\tempytwo-\tempyone)}
	\xdef\myyshifttop{\diffy}
	\end{tikzpicture}
}

\def\myyshiftbot{0}
\def\mypagebot{0}

\newcommand{\marparbot}[1]% content, color
{   \begin{tikzpicture}[remember picture, overlay]
	\ifthenelse{\thepage=\mypagebot}{}{\xdef\myyshiftbot{0}}        
	\xdef\mypagebot{\thepage}
	\node[above right, yshift=\myyshiftbot, text width=\marginparwidth-4pt, inner sep=2pt] (tempnode) at (current page marginpar area.south west) {#1};
	\path (current page marginpar area.south west);
	\pgfgetlastxy{\tempxone}{\tempyone}
	\path (tempnode.north west);
	\pgfgetlastxy{\tempxtwo}{\tempytwo}
	\pgfmathsetmacro{\diffy}{(\tempytwo-\tempyone)}
	\xdef\myyshiftbot{\diffy}
	\end{tikzpicture}
}

\signature{Max Callaghan}

\begin{document}

	\begin{letter}{Adam Yeeles \\ Editor \\ \textit{Nature Climate Change}}

		\begin{tikzpicture}[remember picture, overlay]
		\node [anchor=north west, shift={(-1mm,0mm)}]  at (current page text area.north west) {\footnotesize\textbf{MCC} Torgauer Str. 12-15 \textbar\ 10829 Berlin \textbar\  Germany
		};
		\end{tikzpicture}

		\begin{tikzpicture}[remember picture, overlay]
		\node [anchor=south, shift={(-1mm,5mm)}]  at (current page text area.north east) {\includegraphics[width=4cm]{../pres/MCC_Logo_RZ_rgb.jpg}
		};
		\end{tikzpicture}

		\marpartop{\textcolor[HTML]{0067a1}{\footnotesize\textsf{Mercator Research Institute on Global Commons and Climate Change (MCC) gemeinnützige GmbH}\\ \bigskip}%
			\par\textsf{\textbf{Max W. Callaghan}}%
			\par\textsf{PhD Candidate}%
			\par\textsf{Applied Sustainability Science}%
			\par\textsf{Torgauer Str. 12-15}
			\par\textsf{10829 Berlin | Germany}
			\par\textsf{tel +49 (0) 30 338 55 27 - 245}
			\par\textsf{fax +49 (0) 30 338 55 37 - 102}
			\par\textsf{mail callaghan@mcc-berlin.net}
			\par\textsf{web \url{www.mcc-berlin.net}}
		}

		\marparbot{
			\footnotesize
			Director \\
			Prof. Dr. Ottmar Edenhofer \\
			\medskip
			Bank Account \\
			Sparkasse Berlin \\
			iban DE58 1005 0000 0190 1246 36 \\
			bic  BELADEBEXXX \\
			\medskip
			Amtsgericht Charlottenburg \\
			HRB 142902 B \\
			\medskip
			MCC was founded jointly by the Stiftung Mercator and the Potsdam Institute for Climate Impact Research
		}

		\date{}

   \opening{\today \\ \\ Dear Adam Yeeles,}

   I herewith resubmit on behalf of my co-authors the Letter ``A Topography of Climate Change Literature'' (NCLIM-19071401).

   The letter made use of topic modelling in order to reassess claims made in the literature about how the IPCC has represented scientific knowledge about climate change. This in the context of Big Literature - the hundreds of thousands of papers published on climate change which it is the IPCC's job to assess.

   We express our gratitude for the handling of our manuscript, and for the comments received from the reviewers. We have edited our manuscript accordingly and attach a point by point response to all comments.
   The comments received were clear and along similar lines. Broadly speaking, we 
   \begin{itemize}
   	\item Improved the clarity of our description of our methods in the main text so that they could be understood by a non-specialist audience.
   	\item Added significant nuance to the interpretation of our results. We have been more explicit about what can and can't be drawn as recommendations from our analysis, given its advantages and drawbacks. This has led in turn to a more interesting discussion about how the methods we propose could be used in further IPCC assessments.
   \end{itemize}

	We are confident that this re-submission addresses all of the concerns raised by the reviewers.

\closing{Yours Sincerely,}

\section{Manuscript details}

\begin{tabular}{l r r}

	 & \textbf{Main text} & \textbf{SI \& Methods} \\
	 \textbf{Words} & 2196 & 1584 \\
	 \textbf{Figures} & 4 & 5 \\
	 \textbf{Tables} & 1 & 1 \\
	 \textbf{References} & 19 & 12

\end{tabular} 
%\bibliography{Mendeley}
%\bibliographystyle{unsrt}

\end{letter}
\end{document}