### 2020-01-31 manuscript/supplementary_information.tex
amld pres

\documentclass{article}
\usepackage[a4paper, total={180mm, 260mm}]{geometry}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage{url}
\usepackage{booktabs}
\usepackage{lineno}
\usepackage{color}
\usepackage{caption}
\usepackage{authblk}

\makeatletter
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
	\begin{flushleft}

		{\huge\textbf{\@title}}

		\bigskip

		{\large\textbf{\@author}}

	\end{flushleft}\egroup
}
\makeatother

% Title
\title{A Topography of Climate Change Research - Supplementary Information}
\author[1,2]{Max Callaghan}
\author[1,2]{Jan Minx}
\author[2]{Piers M. Forster}

\affil[1]{Mercator Research Institute on Global Commons and Climate Change, Torgauer Straße, 10829 Berlin, Germany}
\affil[2]{Priestley International Centre for Climate, University of Leeds, Leeds LS2 9JT, United Kingdom}

\begin{document}
	\maketitle

	\renewcommand\thetable{Supplementary .\arabic{table}}

		\begin{table}[htp]
			\begin{center}
				\begin{tabular}{lrrrrr}
					\toprule
					AR &  1 &   2 &   3 &   4 &   5 \\
					WG &    &     &     &     &     \\
					\midrule
					1  & 8\% & 25\% & 37\% & 47\% & 58\% \\
					2  & 6\% & 12\% & 30\% & 38\% & 47\% \\
					3  & 3\% &  9\% & 15\% & 22\% & 35\% \\
					\bottomrule
				\end{tabular}

				\caption*{Supplementary Table 1: The proportion of citations in each report that could be matched with a document in our query from the Web of Science}
				\label{ipcc-matching}
			\end{center}
		\end{table}

\end{document}